---
title: "Rise and fall of programming languages"
author: "Varun Khanna"
date: "15 October 2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let us find the trends in the programming language world. What languages are growing and which are shrinking so that we can tell which are most worth investing time in?

One excellent source of data on the popularity of programming languages is [Stack Overflow](https://stackoverflow.com/). Each Stack Overflow question has a **tag**, which marks a question describing its topic or technology. For instance, there's a tag for languages like [R](https://www.r-project.org/), [Python](https://www.python.org/) or [Perl](https://perl6.org/) and for packages like [ggplot2](https://ggplot2.tidyverse.org/). 

We'll be working with a dataset with one observation for each tag each year. The dataset includes both the number of questions asked in that tag that year and the total number of questions asked in that year.


### Load the libraries and the data 

```{r load-data, message=FALSE}
library("tidyverse")
library("knitr")
data <- read_csv("by_tag_year.csv")
kable(head(data))
```

### Convert into fraction

This data has one observation for each pair of a tag and a year, showing the number of questions asked in that tag in that year and the total number of questions asked in that year. For instance, there were 54 questions asked about the `.htaccess` tag in 2008, out of a total of 58390 questions in that year. Rather than just count, we're probably interested in a percentage: the fraction of questions that year that have that tag. So let's add that to the table.

```{r fraction}
data <- mutate(data, fraction = number/year_total)
kable(head(data))
```

### Has the R been growing or shrinking?

Let's look at whether the fraction of Stack Overflow questions that are about R has been increasing or decreasing over time.

```{r r_over_time}
r_over_time <- filter(data,tag == "r")
kable(head(r_over_time))
```

### Visualizing change over time

Rather than looking at the results in a table, we often want to create a visualization. Change over time is usually visualized with a line plot.

```{r line-plot-r}
ggplot(r_over_time, mapping = aes(x = year, y = fraction)) + 
  geom_line(size = 1) + labs(x = "Year", y = "Fraction of the total questions asked", color = "Tags") + theme_light()
```

**Based on that graph, it looks like R has been growing pretty fast in the last decade.**

Besides R, two other interesting tags are dplyr and ggplot2, which we've already used in this analysis. They both also have Stack Overflow tags! Instead of just looking at R, let's look at all three tags and their change over time. Is each of those tags increasing as a fraction of overall questions? Are any of them decreasing?

```{r selected_tags}
selected_tags <- c("r","dplyr","ggplot2")
selected_tags_over_time <- filter(data, tag %in% selected_tags)

# Plot tags over time
ggplot(selected_tags_over_time, mapping = aes(x = year, y = fraction, color = tag)) +
  geom_line(size = 1) + labs(x = "Year", y = "Fraction of the total questions asked", color = "Tags") + theme_light()
```

### What are the most asked-about tags?

The dplyr and ggplot2 tags may not have as many questions as R, but we can tell they're both growing quickly as well. We might like to know which tags have the most questions *overall*, not just within a particular year. Right now, we have several rows for every tag, but we'll be combining them into one. That means we want `group_by` and `summarize()`

```{r overall-popularity}
# Find total number of questions for each tag
sorted_tags <- data %>% group_by(tag) %>% 
  summarize(tag_total = sum(number)) %>% 
  arrange(desc(tag_total))

kable(head(sorted_tags))
```

### How have large programming languages changed over time?

We've looked at selected tags like R, ggplot2, and dplyr, and seen that they're each growing. What tags might be *shrinking*? A good place to start is to plot the tags that we just saw that were the most-asked-about of all time, including JavaScript, Java, and C#.

```{r highest-tags}
# Get the six largest tags
highest_tags <- head(sorted_tags$tag)
# Filter for the six largest tags
by_tag_subset <- filter(data, tag %in% highest_tags)
# Plot tags over time on a line plot using color to represent tag
ggplot(by_tag_subset, mapping = aes(x = year, y = fraction, color = tag)) +
  geom_line(size = 1) + labs(x = "Year", y = "Fractions of the total questions asked ", color = "Tags") + theme_light()
```

### Some more tags!

Wow, based on that graph we've seen a lot of changes in what programming languages are most asked about. C# gets fewer questions than it used to, and Python has grown quite impressively.

Now let's check out how three big mobile operating systems (Android, iOS, and Windows Phone) have compared in popularity over time.

```{r mobile-os}
# Get tags of interest
os_tags <- c("android","ios","windows-phone")

#Filter for those tags
os_subset <- filter(data, tag %in% os_tags)

# Plot tags over time on a line plot using color to represent tag
ggplot(os_subset, mapping = aes(x = year, y = fraction, color = tag)) + 
  geom_line(size = 1) + labs(x = "Year", y = "Fraction of the total questions asked", color = "Tags") + theme_light()
```

